The Emergent Simplicity of Galaxy Size

### 2017-09-29 paper/galsize_paper.tex
Revised abstract

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius of galaxies, $\rhalf.$ Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. Smaller galaxies cluster stronger relative to larger galaxies of the same stellar mass, a new result. We show that this clustering signal is largely driven by centrals being larger than satellite galaxies of the same halo mass. We use {\tt Halotools} to build a forward model of galaxy size in which $\rhalf$ is set by the halo virial radius $\rvir$ at the time the halo reaches its peak mass, with scatter in $\rhalf$ that is strongly correlated with the halo scale radius $R_{\rm s}$ at that time. The $\rhalf-$dependent clustering of this model is strikingly similar to our SDSS measurements in magnitude, scale-dependence, and $\mstar-$dependence. Together with the result that stellar mass stripping of satellites has only a mild impact on $\rhalf-$dependent clustering, this suggests that the relative size of centrals and satellites is already in place at the time of satellite infall, and supports the notion that $L_{\ast}$ galaxy and halo profiles co-evolve across many Gyr of cosmic time. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}

>>>>

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius of galaxies, $\rhalf.$ Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. Smaller galaxies cluster stronger relative to larger galaxies of the same stellar mass, a new result. We use {\tt Halotools} to test a collection of forward models of galaxy size to identify the qualitative ingredients needed to reproduce the observed clustering. We show that the $\rhalf-$dependence of galaxy clustering is largely driven by centrals being larger than satellite galaxies of the same halo mass. Models in which $\rhalf$ is determined by stellar mass $\mstar$ exhibit qualitatively discrepant clustering properties from SDSS galaxies. Models where $\rhalf$ is linearly proportional to halo virial radius $\rvir$ at the time of peak halo mass are much more successful, provided that scatter in $\rhalf$ is strongly correlated with the halo scale radius $R_{\rm s}$ at that time. Together with the result that stellar mass stripping of satellites has only a mild impact on $\rhalf-$dependent clustering, this suggests that the relative size of centrals and satellites is already in place at the time of satellite infall, and supports the notion that $L_{\ast}$ galaxy and halo profiles co-evolve across many Gyr of cosmic time. 
%Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}
